---
title: "Lec 4 note"
author: "Lau Wai Fung Raymond"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Load the `tidyverse` package

This section loads the packages we need in this lecture.

```{r}
library(tidyverse)
```

### 1. Import and Export V-dem data

This section loads the VDEM dataset and describe its basic information.

```{r}
d <- read_csv("_DataPublic_/vdem/1984_2022/vdem_1984_2022_external.csv")
```

### 2. Select economic development indicators

First, we look at the identifiers of the data.

```{r}
names(d)
```

```{r}
d |> select(country_name, country_id, year) |>
  distinct()
```

```{r}
# Which countries are in this dataset
d |> select(country_name) |> distinct()
```

```{r}
d |> select(year) |> distinct()
```

Select both the country identifiers, GDP, and GDP per capita.

```{r}
d_gdp <- d |> 
  select(country_name, country_id, year, e_gdp, e_gdppc)
d_gdp
```

### Rename Column to Make Names Informative

```{r}
d_gdp <- d_gdp |> 
  rename("GDP" = "e_gdp", "GDP_per_capita" = "e_gdppc", 
         "Country" ="country_name", "ID" = "country_id", 
         "Year" = "year")
d_gdp
```

### 4. `slice` Rows

```{r}
# Want country-years with highest GDP
d_gdp |> 
  slice_max(order_by = GDP, n = 10)
```

```{r}
# Get countries-years with the lowest GDP
d_gdp |> 
  slice_min(order_by = GDP, n = 10)
```

```{r}
set.seed(52)  #required to make research reproducible
d_gdp |> slice_sample(n = 10)  # Sample 10 observations
```

```{r}
d_gdp |> slice_sample(prop = 0.1)
```

### 5. Subset data by row

```{r}
# Want: 2000-2005 data
d_gdp |> 
  filter(Year >= 2000 & Year <= 2005)
```

```{r}
d_gdp |> filter(Country == 'China')
```

```{r}
# Want: 2000-2005 data from China
d_gdp |> 
  filter(Year >= 2000 & Year <= 2005) |> 
  filter(Country == "China")
```

### 6. Arrange

```{r}
# Want: sort the data by GDP per capita
d_gdp |> arrange(GDP_per_capita)
```

```{r}
d_gdp |> arrange(-GDP_per_capita)
```
